---
title: "Simple bar chart"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Minimal bar chart showing the exit polls results of UK elections.

Library needed for this chart:
```{r}
library(ggplot2)
```

Create data frame with election seats results.
```{r preparation}
df <- cbind.data.frame(parties = c("CON", "LAB", "SNP", "LD", "PC", "GRN", "UKIP", "OTH"),
                       seatsExit = c(314, 266, 34, 14, 3, 1, 0, 18))
```

Create custom colour variable from Jinek's chart:
```{r}
pollColours <- c(CON = "#2481CD",
                 LAB = "#981D02",
                 SNP = "#FBEE38",
                 LD = "#DF7200",
                 PC = "#62B988",
                 GRN = "#287C33",
                 UKIP = "#773B90",
                 OTH = "#9FA29F")
```

Bar chart code with ggplot2:
```{r cars}
chart <- ggplot(df) +
  geom_bar(stat = "identity", aes(x = reorder(parties, -seatsExit), 
                                  y = seatsExit, 
                                  fill = parties)) +
  scale_fill_manual(values = pollColours) +
  xlab("Parties") +
  ylab("Number of seats") +
  theme(legend.position = "none",
        panel.background = element_rect(fill = "#F5F0E5")) + 
  geom_text(aes(x = parties,
                y = seatsExit,
                label = seatsExit),
            colour = "black",
            nudge_y = 10)
```

plot:
```{r}
chart
```

### Bar chart tips:
* When you want to flip the axis. Y-axis = parties and X-axis = seats: `chart + coord_flip()` 
* Overrite the bar aestatics. Remove - (dash) from seats axis to get a descended order again.

