---
title: "Regular Expression in R example"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r import, message=FALSE, warning=FALSE}
library(tidyverse)
library(xlsx)
# Import dataset
data <- read.xlsx("~/Work/Projects/Portfolio/asielzoekerbuurten/opvanglocaties_azielzoekers.xlsx", sheetIndex = 1)
# Convert to tibble
data <- as.tibble(data)
```

The following column needs some cleaning up.
```{r}
# The column that needs some cleaning up
data$Locatie
```

## Regular expression pattern

You usually wouldn't remove these parantheses from this column because they are locations and not cities. Also in this dataset you already have a column with cities: `Plaats`. But just for the sake of practice we will remove the parantheses and its content (and the whitespace between the city name and the parantheses). 

```{r message=FALSE, warning=FALSE}
# Remove the parentheses and its content
gsub(" \\([A-Za-z]*\\)", "", as.character(data$Locatie))
```

```{r}
# Store in new data column
data$Locatie <- gsub(".\\([A-Za-z]*\\)", "", as.character(data$Locatie))
```

When you disect the pattern in the `gsub()` regular expression function you should wncode every character from left to right:

- first you see a white space then
- `\\(` : Because `(` is a meta character you have to 'escape' it by adding two backslashes in front.
- `[A-Za-z]*` : This returns any CAPITAL or lower case letter. The `*` says look for 0 or more letters.
- `\\)` : At last we want to return for the closing parenthesis.

This pattern: `" \\([A-Za-z]*\\)"` will return `(RVB)` but also `(regulier)`. 

## Our new dataset
```{r}
data
```


