---
title: "Bar chart"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First things first, start with preparing your data frame.
```{r}
library(tidyverse)
library(eurostat)
library(countrycode)
library(rvest)
library(ggthemes)

# Query eurostat data set
search <- search_eurostat("road", type = "table")
# Accessing "People killed in road accidents" data set.
df <- get_eurostat("tsdtr420")
# Create a data frame filtered on most recent data (2015)
accidents2015 <- 
  get_eurostat("tsdtr420") %>%
  filter(time == "2015-01-01")
# Convert country codes to country names with the countrycode package
accidents2015$country.name <- countrycode(sourcevar = accidents2015$geo, "eurostat", "country.name")
# Create a data frame with only the needed columns
accidents2015Names <- accidents2015[,c(3,6,5)]
# Edit the very long country name "United Kingdom of Great Britain and Northern Ireland""
accidents2015Names$country.name[27] <- "United Kingdom"

```

Now we have the data frame prepared it's time for plotting a bar chart.

First a horizontal bar chart:
```{r}
accidents2015Names %>%
  ggplot(aes(x = country.name, y = values)) +
  geom_col() +
  xlab(NULL)
```

But more suited is the vertical bar chart because the country names are too long as a label for the x axis. It also helps when we order the countries on the number (values) of accidents.
```{r, message=FALSE, warning=FALSE}
accidents2015Names %>%
ggplot(aes(x = reorder(country.name, values), y = values)) +
  geom_col() +
  coord_flip() +
  xlab(NULL) +
  ylab("accidents")
```

We can already quickly gain some insights but still it's not a fair analysis. The countries with the largest number of accidents have also a much bogger population than the other countries. We therefore have to make the values relative to the amount of people. 

First importing the EU population data from wikipedia with the `rvest` package.
```{r}
scotusURL <- "https://en.wikipedia.org/wiki/Demographics_of_the_European_Union"

temp <- scotusURL %>% 
  read_html %>%
  html_nodes("table")

# Storing the retrieved information in a variable and convert list into data frame.
wikiPop <- html_table(temp[1])
wikiPop <- as.data.frame(wikiPop)
```

Now we need the `Population` and the `Member.State` column to merge with the accidents2015 data frame.
```{r}
PopulationEU <- select(wikiPop, Member.State, Population)
# Change the column name to match with the accidents2015 variable.
colnames(PopulationEU) <- c("country.name", "Population")
# Merge data frames
merged <- merge(accidents2015Names, PopulationEU, by = "country.name", all.x = TRUE)
# Remove columns that aren't EU member states.
newMerged <- merged %>%
  filter(country.name != c("Iceland", "Norway"),
         country.name != "Switzerland")
# Convert character to numeric.
newMerged$Population <- gsub(",", "", newMerged$Population)
newMerged$Population <- as.numeric(newMerged$Population)
# Calculate relative number 1:100,000
newMerged <- mutate(newMerged,
       per_100.000 = round((values / Population) * 100000, digits = 1)) %>%
  select(c(1:5))
```

Let's create a more 'honest' representation of the accidents data.
```{r}
newMerged %>%
  ggplot(aes(x = reorder(country.name, per_100.000),
             y = per_100.000)) +
  geom_col() +
  coord_flip() +
  labs(y = "Accidents per 100,000 citizens", # Remember that coord_flip() function is used.
       x = NULL) +
  theme_fivethirtyeight()
```

Now you see that Huge countries like United Kingdom, Germany and Spain are actually doing pretty good.


