---
title: "Building a Bar chart"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this file you'll be guide through the steps of building the famous and renowned bar chart. We'll create horizontal, vertical and reordered charts on values and therefore gain better insights. It's always the case that we import, prepare and tidy our data first before analysising and visualising.

First things first, start with preparing our data frame.
```{r data prep, message=FALSE, warning=FALSE}
library(tidyverse)
library(eurostat)
library(countrycode)
library(rvest)
library(ggthemes)

# Query eurostat data set
search <- search_eurostat("road", type = "table")
# Accessing "People killed in road accidents" data set.
df <- get_eurostat("tsdtr420")
# Create a data frame filtered on most recent data (2015)
accidents2015 <- 
  get_eurostat("tsdtr420") %>%
  filter(time == "2015-01-01")
# Convert country codes to country names with the countrycode package
accidents2015$country.name <- countrycode(sourcevar = accidents2015$geo, "eurostat", "country.name")
# Create a data frame with only the needed columns
accidents2015Names <- accidents2015[,c(3,6,5)]
# Edit the very long country name "United Kingdom of Great Britain and Northern Ireland""
accidents2015Names$country.name[27] <- "United Kingdom"

```

When quickly analysing the data we see that Germany and France have exactly the same number of people killed in road accidents. Normally we should check how the data is gathered etc. but here our main purpose is to visualise the data anylysis. 

Now we have the data frame prepared it's time for plotting a bar chart.

First a horizontal bar chart:
```{r chart1, fig.align = "center"}
accidents2015Names %>%
  ggplot(aes(x = country.name, y = values)) +
  geom_col() +
  xlab(NULL)
```

But more suited is the vertical bar chart because the country names are too long as a label for the x axis. It also helps when we order the countries on the number (values) of accidents.
```{r chart2, fig.align = "center"}
accidents2015Names %>%
ggplot(aes(x = reorder(country.name, values), y = values)) +
  geom_col() +
  coord_flip() +
  labs(x = NULL,
       y = "Deadly accidents",
       title = "People killed in road accidents")
```

We can already quickly gain some insights but still it's not a fair analysis. The countries with the largest number of accidents have also a much larger population than the other countries. We therefore have to make the values relative to the number of people in the each country. 

First importing the EU population data from wikipedia with the `rvest` package.
```{r scraping, message=FALSE, warning=FALSE}
scotusURL <- "https://en.wikipedia.org/wiki/Demographics_of_the_European_Union"

temp <- scotusURL %>% 
  read_html %>%
  html_nodes("table")

# Storing the retrieved information in a variable and convert list into data frame.
wikiPop <- html_table(temp[1])
wikiPop <- as.data.frame(wikiPop)
```

Now we need the `Population` and the `Member.State` column to merge with the accidents2015 data frame.
```{r data prep2, message=FALSE, warning=FALSE}
PopulationEU <- select(wikiPop, Member.State, Population)
# Change the column name to match with the accidents2015 variable.
colnames(PopulationEU) <- c("country.name", "Population")
# Merge data frames
merged <- merge(accidents2015Names, PopulationEU, by = "country.name", all.x = TRUE)
# Remove columns that aren't EU member states.
newMerged <- merged %>%
  filter(country.name != c("Iceland", "Norway"),
         country.name != "Switzerland")
# Convert character to numeric.
newMerged$Population <- gsub(",", "", newMerged$Population)
newMerged$Population <- as.numeric(newMerged$Population)
# Calculate relative number 1:100,000
newMerged <- mutate(newMerged,
       per_100.000 = round((values / Population) * 100000, digits = 1)) %>%
  select(c(1:5))
```

Let's create a more 'honest' representation of the accidents data.
```{r chart3, fig.align = "center"}
newMerged %>%
  ggplot(aes(x = reorder(country.name, per_100.000),
             y = per_100.000)) +
  geom_col(fill = "#B71C1C") +
  coord_flip() +
  labs(x = NULL,
       y = "Deadly accidents",
       title = "People killed in road accidents") +
  theme_fivethirtyeight()
```

Now you see that Huge countries like United Kingdom, Germany and Spain are actually doing pretty good, while countries in the eastern part of EU top the list.


